---
title: "BCC traffic data - exploratory data analysis 01"
author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    # css: custom.css
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    theme: united 
    highlight: pygments 
---

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE, fig.width=8, fig.height=6, dpi=300, out.width="800px", out.height="600px")

# knitr::opts_knit$set(root.dir = 'R:/POP247-Q0786/')
# setwd('R:/POP247-Q0786/')

set.seed(12345)

library(pacman) 

p_load(tidyverse, here, readxl, magrittr, anytime, jsonlite, fs, 
       skimr, statar, naniar, janitor, kableExtra, 
       sf, tmap, tmaptools)

tmap_mode("view") # makes map interactive

isUnique <- function(vector){
  return(!any(duplicated(vector)))
}
```


```{r, echo=FALSE, include=FALSE}
intersect <- jsonlite::fromJSON(here("data", "BCC", "traffic", "raw", "locations",
                                     "Intersection_locations_reference.json"), 
                                flatten = TRUE)
```


```{r, eval=FALSE, include=FALSE}

names(intersect)
dim(intersect)

tab(intersect, scatsVersion)
intersect$scatsVersion <- NULL

tab(intersect, suburb)

isUnique(intersect$subsystem)  #missings!
isUnique(intersect$tsc)
isUnique(intersect$areaNum)

intersect %>%
  select(tsc, coordinates.latLng.latitude, coordinates.latLng.longitude) %>%
  filter(!is.na(coordinates.latLng.longitude) & !is.na(coordinates.latLng.latitude)) %>%
  st_as_sf(., coords = c("coordinates.latLng.longitude", "coordinates.latLng.latitude"),
           crs = 4326, agr = "constant") -> coordinates_sf

qtm(coordinates_sf)

intersect %>% 
  filter(tsc == 8074) %>% 
  View()

```

# Intersection volume 

## Example

```{r}
example <- jsonlite::fromJSON(here("data", "BCC", "traffic", "raw", "temp_02",
                                   "traffic-data-at-int-201809060626.json"), 
                              flatten = TRUE)

```

### Structure

```{r, results='asis'}
skim(example) %>% 
  skimr::kable()
```

### Missing data

```{r}
gg_miss_var(example, show_pct = TRUE)
```


## Larger example 

```{r, include=FALSE, eval=FALSE}
# both solutions need files to be correct jsons! 

# old school - results go to separate files!
names <- list.files(path = here("data", "BCC", "traffic", "raw", "temp_02"), 
                    pattern  =".json") #, full.names = TRUE)

for (i in 1:length(names)) assign(names[i], jsonlite::fromJSON(names[i], flatten=TRUE))

# new school
files <- fs::dir_ls(here("data", "BCC", "traffic", "raw", "temp_02"), regexp = "\\.json$")

example <- map_df(files, jsonlite::fromJSON, flatten = TRUE, .id = "source")

basename(files[1]) # use to clean file names?

# length(files)
# sum_up(file.info(files), size)

```

```{r, echo=FALSE}
```

```{r, include=FALSE, eval=FALSE}
example <- here("data", "BCC", "traffic", "raw", "temp_02") %>% 
  fs::dir_ls(regexp = "\\.json$") %>% 
  map_df(jsonlite::fromJSON, flatten = TRUE, .id = "source")

saveRDS(example,  here("data", "BCC", "traffic", "clean", "example.Rds"))
```

```{r}
example <- readRDS(here("data", "BCC", "traffic", "clean", "example.Rds"))
```

### Time

```{r}
example %<>% 
  select(-source) %>% 
  mutate(recorded = anytime(recorded))

summary(example$recorded)
```

**TODO: identify the gaps**

**TODO: find largest period without data**

**TODO: fill in the gaps**


### Duplicated rows

There are `r nrow(get_dupes(example))` duplicated rows in the raw data - all occurences were removed
```{r}
# isUnique(example$dbid)

example %<>% 
  distinct()
```


### Temporal pattern

#### One intersection

Example from one intersection in St Lucia (`tsc` == 8074). 

Measured flow (`mf`) aggregated for all lanes.

```{r}
traffic_8074 <- example %>% 
  filter(tsc == 8074) %>% 
  mutate(ds = rowSums(select(., contains("ds")), na.rm = TRUE)) %>% 
  mutate(mf = rowSums(select(., contains("mf")), na.rm = TRUE)) %>% 
  mutate(rf = rowSums(select(., contains("rf")), na.rm = TRUE)) %>% 
  group_by(recorded, tsc) %>% 
  summarise(ds = sum(ds),
            mf = sum(mf),
            rf = sum(rf))
``` 

```{r, echo=FALSE}
ggplot(traffic_8074, aes(x = recorded, y = mf)) + 
  geom_line() + scale_x_datetime() + theme_minimal() + 
  xlab("") + ylab("Measured flow (all lanes)")
```

#### All intersections

All (ie. having lat long) across 2 days:

```{r, message=FALSE}
traffic_agg <- example %>% 
  mutate(ds = rowSums(select(., contains("ds")), na.rm = TRUE)) %>% 
  mutate(mf = rowSums(select(., contains("mf")), na.rm = TRUE)) %>% 
  mutate(rf = rowSums(select(., contains("rf")), na.rm = TRUE)) %>% 
  group_by(recorded, tsc) %>% 
  summarise(ds = sum(ds),
            mf = sum(mf),
            rf = sum(rf)) %>% 
  ungroup()

traffic_agg <- left_join(traffic_agg, 
                         select(intersect, tsc, coordinates.latLng.latitude, coordinates.latLng.longitude)) %>% 
  filter(!is.na(coordinates.latLng.longitude) & !is.na(coordinates.latLng.latitude))
```

```{r, echo=FALSE}
ggplot(traffic_agg, aes(x = recorded, y = mf)) + 
  geom_line(aes(group = tsc), alpha = 0.01) + 
  geom_smooth() +
  scale_x_datetime() + theme_minimal() + xlab("") + ylab("Measured flow (all lanes)")
```

#### Hourly aggregates

```{r, echo=FALSE}
traffic_agg %>% 
  mutate(hod = strftime(recorded, format="%H")) %>% 
  ggplot(., aes(x = hod, y = mf)) + 
  geom_boxplot() + theme_minimal() + xlab("Hour of the day") + ylab("Measured flow (all lanes)")
```

#### Daily aggregates

```{r, echo=FALSE}
traffic_agg %>% 
  mutate(dow = strftime(recorded, format="%a")) %>% 
  ggplot(., aes(x = dow, y = mf)) + 
  geom_boxplot() + theme_minimal() + xlab("Day of the week") + ylab("Measured flow (all lanes)")
```

### Spatial pattern

#### One time period

Traffic aggregates across city at 8 AM:

```{r}
traffic_agg %>% 
  filter(recorded == "2018-09-06 08:00:00") %>% 
  ggplot() + 
  geom_point(aes(x=coordinates.latLng.longitude, y=coordinates.latLng.latitude, size=mf), 
             colour = "darkgreen", alpha = 0.5, shape=20, stroke=FALSE) +
  scale_size_continuous(name="Measured flow", range=c(1, 12)) +
  theme_void() + coord_map() 
```

#### Full time period

Traffic aggregates across city for one day:

```{r, eval=FALSE}
library(gganimate)

traffic_agg %>% 
  ggplot() + 
  geom_point(aes(x=coordinates.latLng.longitude, y=coordinates.latLng.latitude, size=mf), 
             colour = "darkgreen", alpha = 0.5, shape=20, stroke=FALSE) +
  scale_size_continuous(name="Measured flow", range=c(1, 12)) +
  theme_void() + coord_map() +
  labs(title = 'Time: {frame_time}') +
  transition_time(recorded)
```

# Session info

```{r}
sessionInfo()
```











